---
title: "V-transform paper: part 2"
author: "Alexander J. McNeil"
date: "2021-01-05"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(knitr.kable.NA = '')
```

This markdown script is the second of three scripts reproducing analyses in the paper "Modelling volatile time series using v-transforms and copulas" by Alexander J.~McNeil (2021).

## Libraries

We need the following:

```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(tscopula) # main library
library(stats4) # for AIC
library(xts) # for handling the time series data
library(report) # for citations of reports
library(rugarch) # comparison of VaR estimate with GARCH
setMethod("logLik", "uGARCHfit", function(object) {
  ll <- likelihood(object)
  attr(ll, "nobs") <- object@model$modeldata$T
  attr(ll, "df") <- length(coef(object))
  class(ll) <- "logLik"
  ll
})
```
## Empirical v-transform

We will first look at the empirical v-transform for data on daily Bitcoin log-returns for the period 2016-2019.

```{r, fig.show='hold', dev.args =list(pointsize=9)}
load("BTCUSD1619.RData")
BTCUSD <- BTCUSD1619$`BTCUSD=X.Close` # closing values
X <- (diff(log(BTCUSD))[-1]) * 100
U <- strank(X)
V <- strank(abs(X))
plot(as.numeric(U), as.numeric(V), xlab= "u", ylab = "v", type="p")
plot(acf(qnorm(V), plot = FALSE), ylab="ACF of z", main= " ")
```

## Fitting VT-ARMA copula models to data

We use the `profilefulcrum` function to investigate different values for the fulcrum.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}

modspec1 <- vtscopula(armacopula(list(ar = 0.25)), Vlinear())
modspec2 <- vtscopula(armacopula(list(ar = 0.95, ma = -0.85)), Vlinear())
modspec3 <- vtscopula(armacopula(list(ar = 0.95, ma = -0.85)), V2p())
modspec4 <- vtscopula(armacopula(list(ar = 0.95, ma = -0.85)), V3p())

deltavals <- c(0, 1/11, 2/11, 3/11, seq(from = 4/11, to = 6/11, length = 50), 7/11, 8/11, 9/11, 1)

tmp1 <- profilefulcrum(U, modspec1, locations = deltavals, plot = FALSE)
tmp2 <- profilefulcrum(U, modspec2, locations = deltavals, plot = FALSE)
tmp3 <- profilefulcrum(U, modspec3, locations = deltavals, plot = FALSE)
tmp4 <- profilefulcrum(U, modspec4, locations = deltavals)

lines(tmp1[,1], tmp1[,2], col =2)
lines(tmp2[,1], tmp2[,2], col =3)
lines(tmp3[,1], tmp3[,2], col =4)
```

We obtain the optimal values of the fulcrum for each v-transform and fit the corresponding models.

```{R, echo = TRUE}
delta4 <- tmp4[tmp4[,2] == max(tmp4[,2]), 1]
delta3 <- tmp3[tmp3[,2] == max(tmp3[,2]), 1]
delta2 <- tmp2[tmp2[,2] == max(tmp2[,2]), 1]
delta1 <- tmp1[tmp1[,2] == max(tmp1[,2]), 1]

modspec1@Vtransform@pars[1] <- as.numeric(delta1)
modspec2@Vtransform@pars[1] <- as.numeric(delta2)
modspec3@Vtransform@pars[1] <- as.numeric(delta3)
modspec4@Vtransform@pars[1] <- as.numeric(delta4)

tsoptions <- list(hessian = TRUE, method = "Nelder-Mead")
vt1arma10 <- fit(modspec1, U, tsoptions = tsoptions)
vt1arma11 <- fit(modspec2, U, tsoptions = tsoptions)
vt2arma11 <- fit(modspec3, U, tsoptions = tsoptions)
vt3arma11 <- fit(modspec4, U, tsoptions = tsoptions)

knitr::kable(AIC(vt1arma10, vt1arma11, vt2arma11, vt3arma11), digits=3)
vt1arma11
```

The model with linear v-transform is favoured (very slightly) by AIC. We carry out a Shapiro test on the residuals and make a number of plots.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
vtcop <- vt1arma11

copmod <- vtcop@tscopula
V <- vtrans(copmod@Vtransform, U)
resid <- kfilter(copmod@Vcopula, V)[, "resid"]
shapiro.test(as.numeric(resid))$p.value

plot(vtcop, plottype="residual")
res <- resid(vtcop)
acf(res)
acf(abs(res))
mu_t <- resid(vtcop, trace = TRUE)
ts.plot(mu_t)
```

## Adding a parametric marginal distribution

We now add parametric margins to the best VT-ARMA(1,1) copula model. 

```{r, fullmodels}
marg_st <- fit(margin("st"), X)
marg_laplace <- fit(margin("laplace", 
                           pars = c(mu = 0.2, scale = 1.5)), X)
marg_dweibull <- fit(margin("doubleweibull", 
                            pars = c(mu = 0.2, shape =0.8, scale = 2.7)), X)

mod_st <- tscm(vtcop, margin = marg_st)
mod_st <- fit(mod_st, X, method = "full", tsoptions = tsoptions)

mod_laplace <- tscm(vtcop, margin = marg_laplace)
mod_laplace <- fit(mod_laplace, X, method = "full", tsoptions = tsoptions)

mod_dweibull <- tscm(vtcop, margin = marg_dweibull)
mod_dweibull <- fit(mod_dweibull, as.numeric(X), method = "full", tsoptions = tsoptions)

knitr::kable(AIC(mod_st, mod_laplace, mod_dweibull), digits = 3)
```

Graphical analyses of fitted double Weibull.

```{R, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
names(mod_dweibull@margin@pars) <- c("mu", "shape","scale")
plot(mod_dweibull, plottype = "margin")
plot(mod_dweibull, plottype = "volprofile")
plot(mod_dweibull, plottype = "vtransform")
plot(mod_dweibull, plottype ="volproxy")
```

Some implications of the fitted model: we estimate the change point value $\mu_T$; we also calculate the volatility response for log-return values of -10 and +10.

```{R, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
qmarg(mod_dweibull@margin, mod_dweibull@tscopula@Vtransform@pars[1])
qnorm(vtrans(mod_dweibull@tscopula@Vtransform,pmarg(mod_dweibull@margin, c(-10,10))))
```

## Comparison with GARCH models

We now compare the VT-ARMA models with GARCH(1,1) models using Student and GED (generalized error distribution) innovations. Note that skewed innovations do not significantly improve the fit of either GARCH model.

```{R, echo = TRUE}
garchmod_ged <- ugarchspec(variance.model = list(
  model="sGARCH", 
  garchOrder=c(1,1)),
  mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
  distribution.model = "ged")
mod_garch_ged <- ugarchfit(garchmod_ged, X)

garchmod_std <- ugarchspec(variance.model = list(
  model="sGARCH", 
  garchOrder=c(1,1)),
  mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
  distribution.model = "std")
mod_garch_std <- ugarchfit(garchmod_std, X)

knitr::kable(AIC(mod_st, mod_laplace, mod_dweibull, 
    mod_garch_std, mod_garch_ged), digits=3)
```
## VaR calculation

The plot shows a comparison of the 95\% VaR estimated by the best VT-ARMA model (red) and the best GARCH model (black).

```{R, echo=TRUE, fig.show='hold', fig.width = 6, fig.height = 4, dev.args =list(pointsize=9)}
alpha <- 0.05
VaR <- quantile(mod_dweibull, alpha)
VaR_alt <- quantile(mod_garch_ged, alpha)
plot.zoo(cbind(X, VaR_alt, VaR), 
         plot.type="single", 
         col=c("grey","black","red"), 
         xlab="", ylab="", type=c("h", "l", "l"))
```


```{R}
cite_packages(sessionInfo())
```
