---
title: "V-transform paper: part 1"
author: "Alexander J. McNeil"
date: "2021-01-05"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(knitr.kable.NA = '')
```

This markdown script is the first of three scripts reproducing analyses in the paper "Modelling volatile time series using v-transforms and copulas" by Alexander J.~McNeil (2020).

## Libraries

We need the following:

```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(tscopula) # main library
library(pBrackets) # add brackets to a plot
library(report) # for citations of reports
```


## A symmetric VT-ARMA process

We specifiy a t distribution for the margin, a symmetric v-transform, a VT-ARMA(1,1) copula process and a VT-ARMA(1,1) process in the following way:

```{r, echo = TRUE}
margmod <- margin("st", pars = c(df = 3, mu = 0, sigma = 1))
vt <- Vsymmetric()
cop_arma <- armacopula(list(ar = 0.95, ma = -0.85))
cop_vtarma <- vtscopula(cop_arma, Vtransform = vt)
vtarma11 <- tscm(cop_vtarma, margmod)
```

We will generate realizations of the $(X_t)$, $(U_t)$, $(V_t)$ and $(Z_t)$ processes.

```{r, echo = TRUE}
set.seed(199)
Xsim <- sim(vtarma11, n = 500) 
Usim <- pmarg(margmod, Xsim)
Vsim <- vtrans(vt, Usim)
Zsim <- qnorm(Vsim)
```


Now we plot $(X_t)$ and $(Z_t)$ along with the acf of $(X_t)$ and $(|X_t|)$.
```{r, fig.show='hold', dev.args =list(pointsize=9)}
ts.plot(Xsim, type="h", ylab=expression(X[t]))
ts.plot(Zsim, type="h", ylab=expression(Z[t]))
plot(acf(Xsim, plot = FALSE), xlab="k", 
     ylab = expression(rho(X[t],X[t+k])), main="", cex =0.5)
plot(acf(abs(Xsim), plot = FALSE), xlab="k", 
     ylab = expression(rho(abs(X[t]),abs(X[t+k]))), main="")
```


## Visualizing a v-transform

The following code gives a picture of a 3-parameter v-transform and annotates it to illustrate the square property and show how stochastic inversion works.

```{r, echo=TRUE, fig.width = 4.7, fig.height = 5, fig.show='hold', dev.args =list(pointsize=9)}
V <- V3p(delta = 0.55, kappa = 1.4, xi = 0.65)
plot(V)
upt <- 0.285
vpt <- vtrans(V, upt)
udual <- vinverse(V, vpt) + vpt
abline(h = vpt, lty =2)
lines(c(upt,upt), c(-0.05, vpt), lty=2)
lines(c(udual, udual), c(-0.05, vpt), lty=2)
axis(2, at = vpt, "v")
axis(1, at = upt, expression(u))
axis(1, at = udual, expression(paste(u,"*")))
axis(1, at = 0.55, expression(delta))
brackets(upt, vpt+0.01, udual, vpt+0.01, type=1)
text(upt+vpt/2, vpt+0.075, "v")
```

## A model with asymmetric v-transform

This time we create and simulate a model with an asymmetric v-transform.

```{r, echo=TRUE}
margmod <- margin("sst", c(df=3, gamma=0.8, mu=0.3, sigma=1.0))
downprob <- pmarg(margmod, 0)
cop_arma <- armacopula(list(ar = 0.95, ma = -0.85))
vt <- V3p(delta = downprob, kappa = 0.9, xi = 1.1)
cop_vtarma <- vtscopula(cop_arma, Vtransform = vt)
vtarma11 <- tscm(cop_vtarma, margmod)
```


```{r, echo=TRUE}
n <- 500
set.seed(9)
Xsim <- sim(vtarma11, n)
Usim <- pmarg(margmod, Xsim)
Vsim <- vtrans(vt, Usim)
```

We plot both $(X_t)$ and $(Z_t)$ (in red). On the latter we superimpose $(\mu_t)$ calculated using the Kalman filter and draw lines to indicate high and low levels of the conditional mean.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
ts.plot(Xsim, type="h", ylab=expression(X[t]))
ts.plot(qnorm(Vsim), type="h", ylab=expression(Z[t]), col = 2)
kf <- kfilter(cop_arma, Vsim)
lines(kf[,"mu_t"])
mu_t_1 <- 0.5
mu_t_2 <- -0.5
abline(h=mu_t_1)
abline(h=mu_t_2)
```

Let's look at the conditional densities when $\mu_t$ takes the values indicated by the two horizontal lines. The left picture shows a high volatility situation (bimodal) and the right picture a low volatility situation.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9), results='asis'}
x <- seq(from=-7, to=5, length=1000)
sigmaeps <- sigmastarma(cop_arma)
y1 <- tscopula:::dcondvtarma(x, vtarma11, mu_t_1, sigmaeps)
y2 <- tscopula:::dcondvtarma(x, vtarma11, mu_t_2, sigmaeps)
y0 <- tscopula:::dcondvtarma(x, vtarma11, 0, 1)
plot(x, y1, type="l", ylim = range(y1, y0), ylab = "density")
lines(x, y0, lty=2)
plot(x, y2, type="l", ylim = range(y2, y0), ylab = "density")
lines(x, y0, lty=2)
```


```{R}
cite_packages(sessionInfo())
```
