---
title: "V-transform paper: part 3"
author: "Alexander J. McNeil"
date: "2021-01-05"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
options(knitr.kable.NA = '')
```

This markdown script is the third of three scripts reproducing analyses in the paper "Modelling volatile time series using v-transforms and copulas" by Alexander J.~McNeil (2021).

## Libraries

We need the following:

```{r, echo = TRUE, warning = FALSE, message = FALSE}
library(tscopula) # main library
library(stats4) # for AIC
library(xts) # for handling the time series data
library(report) # for citations of reports
library(rugarch) # comparison of VaR estimate with GARCH
setMethod("logLik", "uGARCHfit", function(object) {
  ll <- likelihood(object)
  attr(ll, "nobs") <- object@model$modeldata$T
  attr(ll, "df") <- length(coef(object))
  class(ll) <- "logLik"
  ll
})
```

## Data

The model was fitted to the Bitcoin log-returns for the period 2016-2019.

```{r}
load("BTCUSD1619.RData")
BTCUSD <- BTCUSD1619$`BTCUSD=X.Close` # closing values
X <- (diff(log(BTCUSD))[-1]) * 100
U <- strank(X)
```


## Best fitting ARMA Copula Model

We construct again the best model.

```{r, echo=TRUE, fig.show='hold', dev.args =list(pointsize=9)}
tsoptions <- list(hessian = TRUE, method = "Nelder-Mead")
modspec <- vtscopula(armacopula(list(ar = 0.95, ma = -0.85)), Vlinear(0.460113))
vtcop <- fit(modspec, U)
marg_dweibull <- fit(margin("doubleweibull", 
                            pars = c(mu = 0.2, shape =0.8, scale = 2.7)), X)
mod_dweibull <- tscm(vtcop, margin = marg_dweibull)
mod_dweibull <- fit(mod_dweibull, as.numeric(X), method = "full", tsoptions = tsoptions)
AIC(mod_dweibull)
```

## Out-of-sample forecasting

This analysis is rather slow. We first set up the data for forecasting.

```{R, echo = TRUE}
load("BTCUSD1219.RData")
BTCUSD1219 <- BTCUSD1219$`BTCUSD=X.Close` # closing values
X2 <- (diff(log(BTCUSD1219))[-1]) * 100
X2 <- X2[-(1:43)]
n0 <- 1000 # length of moving window
nforecast <- length(X2) - n0 # number of forecasts (period 2016-19)
```

We use the best VT-ARMA model for forecasting.

```{R, echo = TRUE}
tsoptions <- list(hessian = FALSE, method = "Nelder-Mead")
VaR <- matrix(NA,ncol=2, nrow = nforecast)
for (i in 1:nforecast){
  fitdata <- as.numeric(X2[i:(i+n0-1)])
  fit <- fit(mod_dweibull, fitdata, tsoptions = tsoptions)
  VaR[i,1] <- quantile(fit, 0.05,last = TRUE)
  VaR[i,2] <- quantile(fit, 0.01, last = TRUE)
}
```

We also use the best GARCH model for forecasting.

```{R, echo = TRUE}
garchmod_ged <- ugarchspec(variance.model = list(
  model="sGARCH", 
  garchOrder=c(1,1)),
  mean.model=list(armaOrder=c(0,0),include.mean=TRUE),
  distribution.model = "ged")
roll <- ugarchroll(garchmod_ged, X2, n.ahead = 1, 
                   forecast.length = nforecast, 
                   refit.every = 1, 
                   refit.window = "moving",
                   VaR.alpha = c(0.05,0.01), keep.coef = F)
```

We compare results and plot them as before.

```{R, echo = TRUE, fig.show='hold', fig.width = 6, fig.height = 4, dev.args =list(pointsize=9)}
sum(roll@forecast$VaR$realized < roll@forecast$VaR$`alpha(5%)`)
sum(roll@forecast$VaR$realized < VaR[,1])
nforecast*0.05

sum(roll@forecast$VaR$realized < roll@forecast$VaR$`alpha(1%)`)
sum(roll@forecast$VaR$realized < VaR[,2])
nforecast*0.01

Xforecast <- X2[-(1:n0)]
VaR_GARCH <- roll@forecast$VaR$`alpha(5%)`
attributes(VaR_GARCH) <- attributes(Xforecast)
VaR_VTARMA <- VaR[,1]
attributes(VaR_VTARMA) <- attributes(Xforecast)
plot.zoo(cbind(Xforecast, VaR_GARCH, VaR_VTARMA), 
         plot.type="single", 
         col=c("grey","black","red"), 
         xlab="", ylab="", type=c("h", "l", "l"))
```


```{R}
cite_packages(sessionInfo())
```
